Appendix: Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-30 appendix.qmd
Add different options in our model to Appendix

data <- data |>
  mutate(ID = row_number())

data_split <- initial_split(data)
training_data <- training(data_split)
test_data <- testing(data_split)
```

@@

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

@@

@@

>>>>

training_data <- read_rds("data/s_curve/s_curve_training.rds")
test_data <- read_rds("data/s_curve/s_curve_test.rds")
```

@@

UMAP_data <- read_rds("data/s_curve/s_curve_umap.rds")

@@

## Notations

@@

## Different options to construct the 2D model and lift the model into high-D space

Mainly, the hexagonal bin centroids are used to construct 2D model and averaged high-D data is used to lift the model into high-D space. As options except bin centroids, bin means can used to construct the 2D model. Also, except averaged high-D data, weighted mean can used to lift the model into high-D space. If we use bin centers instead of bin centroids, the triangles are not regular.

- Why bin means? because the distribution of points within hexagonals are not always reprsented by the center.

### Case 1: Construct 2D model with hexbin means and lift the model into high-D space with averaged high-D data

## Import data

UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")

## UMAP

num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
                                         cell_area = 1)
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data

## To generate the grid
hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
## Data set with all possible centroids in the hexagonal grid
full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)

## Triangulate bin centroids
tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)

# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)

## To find the benchmark value
benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
benchmark_umap_s_curve <- 2.0384346

# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)

# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()
# 
# trimesh_umap_s_curve <- trimesh_umap_s_curve +
#   #ggtitle("(a)") +
#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
#   theme_light() +
#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
#   ) +
#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
# theme(axis.text = element_text(size = 5),
#       axis.title = element_text(size = 7))

trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# ### Define type column
# df <- df_all_umap_s_curve |>
#   dplyr::select(tidyselect::starts_with("x")) |>
#   dplyr::mutate(type = "data") ## original dataset
# 
# df_b <- df_bin_umap_s_curve |>
#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
#   dplyr::select(-hb_id) |>
#   dplyr::mutate(type = "model") ## Data with summarized mean
# 
# df_exe <- dplyr::bind_rows(df_b, df)
# 
# distance_df_small_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance < benchmark_umap_s_curve)
# 
# distance_df_long_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance >= benchmark_umap_s_curve)
# 
# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
# 
# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
# 
# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = df_bin_centroids_n, aes(x = x, y = y), color = "#33a02c") +
  geom_point(data = df_bin_centroids_umap_s_curve, aes(x = x, y = y), color = "black")

## UMAP
## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                   x = UMAP1, y = UMAP2)

num_bins_vec <- 1:13 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names

eval_data_training <- dplyr::bind_rows(vec)[0, ]
eval_data_training <- eval_data_training |>
  dplyr::mutate_if(is.character, as.numeric)

for (i in 1:length(num_bins_vec)) {

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_df_training <- eval_df_training |>
    mutate(num_bins_x = num_bins_vec[i])

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

eval_data_training <- eval_data_training |>
  mutate(data_type = "training")

MSE_df_2 <- eval_data_training |>
  dplyr::mutate(method = "UMAP")
```

#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
MSE_df <- MSE_df_2

MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the most effective NLDR technique to represent the original data in a 2D space? What is the effective number of bins in each NLDR technique to create a 2D model?"
#| label: fig-diagnosticpltScurve
##| out-width: 100%

trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

### Case 2: Construct 2D model with hexbin centroid and lift the model into high-D space with weighted mean of high-D data

## Import data

UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")

## UMAP

num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
                                         cell_area = 1)
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data

## Data set with all possible centroids in the hexagonal grid
full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_umap_s_curve)
## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)

## Weighted averaged on high-D
hb_object <- hexbin_data_object_umap_s_curve$hb_data
df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")

## Triangulate bin centroids
tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)

# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)

## To find the benchmark value
benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
#benchmark_umap_s_curve <- 2.0384346

# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)

# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()
# 
# trimesh_umap_s_curve <- trimesh_umap_s_curve +
#   #ggtitle("(a)") +
#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
#   theme_light() +
#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
#   ) +
#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
# theme(axis.text = element_text(size = 5),
#       axis.title = element_text(size = 7))

trimesh_removed_umap_s_curve_bin_centroid <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_removed_umap_s_curve_bin_centroid <- trimesh_removed_umap_s_curve_bin_centroid +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# ### Define type column
# df <- df_all_umap_s_curve |>
#   dplyr::select(tidyselect::starts_with("x")) |>
#   dplyr::mutate(type = "data") ## original dataset
# 
# df_b <- df_bin_umap_s_curve |>
#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
#   dplyr::select(-hb_id) |>
#   dplyr::mutate(type = "model") ## Data with summarized mean
# 
# df_exe <- dplyr::bind_rows(df_b, df)
# 
# distance_df_small_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance < benchmark_umap_s_curve)
# 
# distance_df_long_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance >= benchmark_umap_s_curve)
# 
# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
# 
# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
# 
# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)

## UMAP
## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                   x = UMAP1, y = UMAP2)

num_bins_vec <- 1:13 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names

eval_data_training <- dplyr::bind_rows(vec)[0, ]
eval_data_training <- eval_data_training |>
  dplyr::mutate_if(is.character, as.numeric)

for (i in 1:length(num_bins_vec)) {

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_df_training <- eval_df_training |>
    mutate(num_bins_x = num_bins_vec[i])

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

eval_data_training <- eval_data_training |>
  mutate(data_type = "training")

MSE_df_2 <- eval_data_training |>
  dplyr::mutate(method = "UMAP")
```

#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
MSE_df <- MSE_df_2

MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the most effective NLDR technique to represent the original data in a 2D space? What is the effective number of bins in each NLDR technique to create a 2D model?"
#| label: fig-diagnosticpltScurve2
##| out-width: 100%

trimesh_removed_umap_s_curve_bin_centroid + aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

### Case 3: Construct 2D model with hexbin means and lift the model into high-D space with weighted mean of high-D data

## Import data

UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")

## UMAP

num_bins_umap_s_curve <- calculate_effective_x_bins(.data = UMAP_s_curve, x = UMAP1,
                                         cell_area = 1)
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_mean(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data

## To generate the grid
hexbin_data_object_n <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)
df_bin_centroids_n <- hexbin_data_object_n$hexdf_data
## Data set with all possible centroids in the hexagonal grid
full_centroid_df <- generate_full_grid_centroids(df_bin_centroids_n)
## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)

## Weighted averaged on high-D
hb_object <- hexbin_data_object_umap_s_curve$hb_data
df_bin_umap_s_curve <- weighted_highD_data(training_data, nldr_df_with_id = UMAP_data_with_hb_id_s_curve, hb_object, column_start_text = "x")

## Triangulate bin centroids
tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)

# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)

## To find the benchmark value
benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)
benchmark_umap_s_curve <- 2.0384346

# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)

# trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()
# 
# trimesh_umap_s_curve <- trimesh_umap_s_curve +
#   #ggtitle("(a)") +
#   xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
#   theme_light() +
#   theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
#   ) +
#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
# theme(axis.text = element_text(size = 5),
#       axis.title = element_text(size = 7))

trimesh_removed_umap_s_curve_bin_mean <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_removed_umap_s_curve_bin_mean <- trimesh_removed_umap_s_curve_bin_mean +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# ### Define type column
# df <- df_all_umap_s_curve |>
#   dplyr::select(tidyselect::starts_with("x")) |>
#   dplyr::mutate(type = "data") ## original dataset
# 
# df_b <- df_bin_umap_s_curve |>
#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
#   dplyr::select(-hb_id) |>
#   dplyr::mutate(type = "model") ## Data with summarized mean
# 
# df_exe <- dplyr::bind_rows(df_b, df)
# 
# distance_df_small_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance < benchmark_umap_s_curve)
# 
# distance_df_long_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance >= benchmark_umap_s_curve)
# 
# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
# 
# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
# 
# show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)

## UMAP
## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                   x = UMAP1, y = UMAP2)

num_bins_vec <- 1:13 ## Number of bins along the x-axis

vec <- stats::setNames(rep("", 5), c("number_of_bins", "number_of_observations", "total_error", "total_mse", "num_bins_x"))  ## Define column names

eval_data_training <- dplyr::bind_rows(vec)[0, ]
eval_data_training <- eval_data_training |>
  dplyr::mutate_if(is.character, as.numeric)

for (i in 1:length(num_bins_vec)) {

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_s_curve, nldr_df_test = UMAP_s_curve, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_df_training <- eval_df_training |>
    mutate(num_bins_x = num_bins_vec[i])

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

eval_data_training <- eval_data_training |>
  mutate(data_type = "training")

MSE_df_2 <- eval_data_training |>
  dplyr::mutate(method = "UMAP")
```

#MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)
MSE_df <- MSE_df_2

MSE_df$method <- factor(MSE_df$method, levels = c("UMAP"))

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training") |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df |> dplyr::filter(method == "UMAP"), aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.position = "none", legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_colour_manual(values = c("#377eb8", "#e41a1c", "#4daf4a", "#984ea3", "#ff7f00")) +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the most effective NLDR technique to represent the original data in a 2D space? What is the effective number of bins in each NLDR technique to create a 2D model?"
#| label: fig-diagnosticpltScurve3
##| out-width: 100%

trimesh_removed_umap_s_curve_bin_mean + aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```
